Untitled

### 2021-03-12 ms/manuscript.Rmd
Update MS

The degree of DT in filmy fern sporophytes occupying different niches has been relatively well studied.
In a series of elegant experiments on Jamaican filmy ferns, @Shreve1911 demonstrated that water-use strategies differ between species occupying contrasting habitats (\ie, terrestrial, low elevation epiphyte, and high elevation epiphyte).
Proctor [-@Proctor2003; -@Proctor2012] investigated water relations and the ability to recover from desiccation in sporophytes of several filmy ferns, and found a general correlation between high DT and adaptation for high light levels in widespread species vs. low DT and adaptation to low light levels in species occupying more sheltered habitats.
In the only study to our knowledge including in situ physiological observations of hymenophyllaceous gametophytes, @Johnson2000 found that gametophytes of *Vandenboschia speciosa* (Willd.) G. Kunkel (= *Trichomanes speciosum* Willd.) are adapted for extremely low light environments, and operate at lower light levels than conspecific sporophytes.
@Parra2009 and @Saldana2013 investigated the relationship between physiology (photosynthetic rates and water relations) and vertical stratification in several epiphytic filmy fern species (primarily genus *Hymenophyllum*) along tree trunks in Chile.
They found that species occurring at greater heights in the trees tended to have greater DT, higher photosynthetic capacity (A~max~), and lower rates of evapotranspiration.
A study using two species of *Hymenophyllum* demonstrated that DT in filmy ferns likely operates via a homoiochlorophyllous, constitutive mechanism, which is more similar to that of bryophytes than other drought-tolerant vascular plants, thus allowing filmy ferns to adjust to dry conditions extremely rapidly [@GarcesCea2014].
This was further supported by evidence that induced biomechanical mechanisms likely do not play a role in DT of filmy ferns [@Fallard2018; but see @Mkhize2020].
Recently, transcriptomics has been applied to elucidate the genes involved at each step of the desiccation-recovery cycle of filmy ferns [@Ostria-Gallardo2020; @Ostria-Gallardo2020a]. 
In epiphytic *Hymenoglossum cruentum* Cav. from Chile, key genes and gene networks were identified related to intracellular mobility and photosynthetic metabolism in the desiccated state, and those related to detoxifying pathways and stabilization of photosystems during rehydration [@Ostria-Gallardo2020a].
In a comparison of low-trunk and canopy epiphytes, @Ostria-Gallardo2020 identified multiple differentially expressed genes that may be specific to tolerating desiccation in these microhabitats.

@@

@@

\*\*\* NOTE TO REVIEWERS: The dataset and code will not be made public until the paper is accepted. While the paper is under review, these links can be used for the dataset (https://figshare.com/s/d6349abf01a3756a5aae) and code (https://www.dropbox.com/sh/h788g1h3sjcqj0j/AACwFWTmL6GxrbLkCfOrXZU9a?dl=0)  \*\*\*

>>>>

The degree of DT in filmy fern sporophytes occupying different niches has been relatively well studied.
In a series of elegant experiments on Jamaican filmy ferns, @Shreve1911 demonstrated that water-use strategies differ between species occupying contrasting habitats (\ie, terrestrial, low elevation epiphyte, and high elevation epiphyte).
Proctor [-@Proctor2003; -@Proctor2012] investigated water relations and the ability to recover from desiccation in sporophytes of several filmy ferns, and found a general correlation between high DT and adaptation for high light levels in widespread species vs. low DT and adaptation to low light levels in species occupying more sheltered habitats.
In the only study to our knowledge including in situ physiological observations of hymenophyllaceous gametophytes, @Johnson2000 found that gametophytes of *Vandenboschia speciosa* (Willd.) G. Kunkel (= *Trichomanes speciosum* Willd.) are adapted for extremely low light environments, and operate at lower light levels than conspecific sporophytes.
@Parra2009 and @Saldana2013 investigated the relationship between physiology (photosynthetic rates and water relations) and vertical stratification in several epiphytic filmy fern species (primarily genus *Hymenophyllum*) along tree trunks in Chile.
They found that species occurring at greater heights in the trees tended to have greater DT, higher photosynthetic capacity (A~max~), and lower rates of evapotranspiration.
A study using two species of *Hymenophyllum* demonstrated that DT in filmy ferns likely operates via a homoiochlorophyllous, constitutive mechanism, which is more similar to that of bryophytes than other drought-tolerant vascular plants, thus allowing filmy ferns to adjust to dry conditions extremely rapidly [@GarcesCea2014].
This was further supported by evidence that induced biomechanical mechanisms likely do not play a role in DT of filmy ferns [@Fallard2018; but see @Mkhize2020].
Recently, transcriptomics has been applied to elucidate the genes involved at each step of the desiccation-recovery cycle of filmy ferns [@Ostria-Gallardo2020; @Ostria-Gallardo2020a]. 
In epiphytic *Hymenoglossum cruentum* Cav., key genes and gene networks were identified related to intracellular mobility and photosynthetic metabolism in the desiccated state, and those related to detoxifying pathways and stabilization of photosystems during rehydration [@Ostria-Gallardo2020a].
In a comparison of low-trunk and canopy epiphytes, @Ostria-Gallardo2020 identified multiple differentially expressed genes that may be specific to tolerating desiccation in these microhabitats.

@@

Fieldwork was done under permits issued by the French Polynesian Government (Délégation à la Recherche) and the Haut-commissariat de la République en Polynésie Francaise (Protocole d’Accueil 2012–2014). 
Voucher specimens were deposited at UC, with duplicates at GH and PAP (abbreviations follow @Thiers2021).

@@

\*\*\* NOTE TO REVIEWERS: The DOI for the dataset will not be active until the paper is accepted. Until then, this link can be used for the dataset https://figshare.com/s/d6349abf01a3756a5aae  \*\*\*
